Lesson 3
========================================================

***

### What to Do First?
Notes:

***

### Pseudo-Facebook User Data
Notes:

```{r Pseudo-Facebook User Data}
setwd("/Users/markmeleka/Documents/Comp Sci and Programming Learning/Udacity/DAND/Exploratory_Data_Analysis/eda-course-materials/lesson3")

pf <- read.csv("pseudo_facebook.tsv", sep = "\t")
names(pf)
```

***

### Histogram of Users' Birthdays
Notes:

```{r Histogram of Users\' Birthdays}
#install.packages('ggplot2')
#library(ggplot2)

qplot(data = pf, x = dob_day) +
  scale_x_continuous(breaks = 1:31)

ggplot(aes(x = dob_day), data = pf) +
  geom_histogram(binwidth = 1) +
  scale_x_continuous(breaks = 1:31)
```

***

#### What are some things that you notice about this histogram?
Response:

***

### Moira's Investigation
Notes:

***

### Estimating Your Audience Size
Notes:

***

#### Think about a time when you posted a specific message or shared a photo on Facebook. What was it?
Response:

#### How many of your friends do you think saw that post?
Response:

#### Think about what percent of your friends on Facebook see any posts or comments that you make in a month. What percent do you think that is?
Response:

***

### Perceived Audience Size
Notes:

***
### Faceting
Notes:

```{r Faceting}
qplot(data = pf, x = dob_day) +
  scale_x_continuous(breaks = 1:31) + 
  facet_wrap(~dob_month, ncol = 4)
```

#### Let’s take another look at our plot. What stands out to you here?
Response:

***

### Be Skeptical - Outliers and Anomalies
Notes:

***

### Moira's Outlier
Notes:
#### Which case do you think applies to Moira’s outlier?
Response:

***

### Friend Count
Notes:

#### What code would you enter to create a histogram of friend counts?

```{r Friend Count}
qplot(data = pf, x = friend_count)
```
```{r Friend Count #2}
ggplot(aes(x = friend_count), data = pf) +
  geom_histogram()
```


#### How is this plot similar to Moira's first plot?
Response: It's right skewed

***

### Limiting the Axes
Notes:

```{r Limiting the Axes}
qplot(data=pf, x= friend_count, xlim = c(0, 1000))

qplot(data=pf, x= friend_count) +
  scale_x_continuous(limits = c(0, 1000))
```

```{r Limiting the Axes}
ggplot(aes(x = friend_count), data = pf) +
  geom_histogram() +
  scale_x_continuous(limits = c(0, 1000))
```

  
### Exploring with Bin Width
Notes:

***

### Adjusting the Bin Width
Notes:

```{r}
qplot(x = friend_count, data = pf, binwidth = 25) +
  scale_x_continuous(limits = c(0, 1000), breaks = seq(0, 1000, 50))

#Equivalent ggplot syntax:

ggplot(aes(x = friend_count), data = pf) +
  geom_histogram(binwidth = 25) +
  scale_x_continuous(limits = c(0, 1000), breaks = seq(0, 1000, 50))
```

### Faceting Friend Count
```{r Faceting Friend Count}
# What code would you add to create a facet the histogram by gender?
# Add it to the code below.
qplot(x = friend_count, data = pf, binwidth = 10) +
  scale_x_continuous(limits = c(0, 1000),
                     breaks = seq(0, 1000, 50)) + 
  facet_wrap(~gender)

#Equivalent ggplot syntax:

ggplot(aes(x = friend_count), data = pf) +
  geom_histogram() +
  scale_x_continuous(limits = c(0, 1000), breaks = seq(0, 1000, 50)) +
  facet_wrap(~gender)
```

***

### Omitting NA Values
Notes:

```{r Omitting NA Values}
qplot(x = friend_count, data = subset(pf, !is.na(gender)), binwidth = 10) +
  scale_x_continuous(limits = c(0, 1000),
                     breaks = seq(0, 1000, 50)) + 
  facet_wrap(~gender)
```

***

### Statistics 'by' Gender
Notes:

```{r Statistics \'by\' Gender}
table(pf$gender)
by(pf$friend_count, pf$gender, summary)
```

#### Who on average has more friends: men or women?
Response:

#### What's the difference between the median friend count for women and men?
Response:

#### Why would the median be a better measure than the mean?
Response:

***

### Tenure
Notes:

```{r Tenure}

qplot(data = pf, x = tenure, color = I('black'), fill = I('#099DD9'), binwidth = 30) + 
  scale_x_continuous(breaks = seq(0,3000, 400))

#Equivalent ggplot syntax for plots:

ggplot(aes(x = tenure), data = pf) +
  geom_histogram(binwidth = 30, color = 'black', fill = '#099DD9')

#by tenure/365
ggplot(aes(x = tenure/365), data = pf) +
  geom_histogram(binwidth = .25, color = 'black', fill = '#F79420')


```

***

#### How would you create a histogram of tenure by year?

```{r Tenure Histogram by Year}

```

***

### Labeling Plots
Notes:

```{r Labeling Plots}
#with labels
ggplot(aes(x = tenure / 365), data = pf) +
  geom_histogram(color = 'black', fill = '#F79420') +
  scale_x_continuous(breaks = seq(1, 7, 1), limits = c(0, 7)) +
  xlab('Number of years using Facebook') +
  ylab('Number of users in sample')
```

***

### User Ages
Notes:

```{r User Ages}
qplot(data = pf, x = age, color= I('black'), fill=I('grey'), binwidth = 2) +
  scale_x_continuous(breaks = seq(0,120,20)) +
  xlab('User Age') +
  ylab('# users in sample')

#Equivalent ggplot syntax:

ggplot(aes(x = age), data = pf) +
  geom_histogram(binwidth = 1, fill = '#5760AB') +
  scale_x_continuous(breaks = seq(0, 113, 5))
```

#### What do you notice?
Response:

***

### The Spread of Memes
Notes:

***

### Lada's Money Bag Meme
Notes:

***

### Transforming Data
Notes:
```{r}
#install.packages("gridExtra")
library('gridExtra')
```


***

### Add a Scaling Layer
Notes:

```{r Add a Scaling Layer}
p1 <- qplot(data = pf, x = friend_count)
p2 <- qplot(data = pf, x = log10(friend_count+1))
p3 <- qplot(data = pf, x = sqrt(friend_count))
grid.arrange(p1,p2,p3, ncol = 1)
```

```{r}
#using ggplot and different method for scaling x
p1 <- ggplot(data = pf, aes(x = friend_count)) + geom_histogram()
p2 <- p1 + scale_x_log10()
p3 <- p1 + scale_x_sqrt()
grid.arrange(p1,p2,p3, ncol = 1)
```

***


### Frequency Polygons

```{r Frequency Polygons}
qplot(x = friend_count, data = subset(pf, !is.na(gender)), binwidth = 10, geom = 'freqpoly', color = gender) +
  scale_x_continuous(limits = c(0, 1000),
                     breaks = seq(0, 1000, 50))
```

```{r Frequency Polygons w/ proportions}
qplot(x = friend_count, y = ..count../sum(..count..),
      data = subset(pf, !is.na(gender)), 
      xlab = 'Friend Count',
      ylab = 'Proportion of users with that friend count',
      binwidth = 10, geom = 'freqpoly', 
      color = gender) +
  scale_x_continuous(limits = c(0, 1000),
                     breaks = seq(0, 1000, 50))
```
***

### Likes on the Web
Notes:

```{r Likes on the Web}
# **What is this doing? Why does adding scale_x_log10 change the graph so much?
qplot(x = www_likes,
      data = subset(pf, !is.na(gender)), 
      geom = 'freqpoly', 
      color = gender) +
  scale_x_continuous(breaks = seq(0,10000)) + 
  scale_x_log10()
```
```{r Likes on the Web w/ ggplot}
ggplot(aes(x = www_likes), data = subset(pf, !is.na(gender))) +
  geom_freqpoly(aes(color = gender)) +
  scale_x_log10()
```

```{r Likes on the Web}
by(pf$www_likes, pf$gender, sum)
```

***

### Box Plots
Notes:

```{r Box Plots}
#this way of limiting the y-axis is incorrect because it calculates the box-whisker plot *after* removing data points
qplot(x=gender, y= friend_count, 
      data = subset(pf, !is.na(gender)), 
      geom='boxplot')
  
```

#### Adjust the code to focus on users who have friend counts between 0 and 1000.

```{r}
#this way of limiting the y-axis is incorrect because it calculates the box-whisker plot *after* removing data points
qplot(x=gender, y= friend_count, 
      data = subset(pf, !is.na(gender)), 
      geom='boxplot', ylim = c(0,1000))

#this way is more accurate because it calculates the box-whisker plot including *all* the data points
qplot(x=gender, y= friend_count, 
      data = subset(pf, !is.na(gender)), 
      geom='boxplot') + 
  coord_cartesian(ylim = c(0,1000))
```

***

### Box Plots, Quartiles, and Friendships
Notes:

```{r Box Plots, Quartiles, and Friendships}

```

#### On average, who initiated more friendships in our sample: men or women?
Response: Women, but not by much.
#### Write about some ways that you can verify your answer.
Response:
```{r Friendships initiated by Gender}
qplot(x=gender, y= friendships_initiated, 
      data = subset(pf, !is.na(gender)), 
      geom='boxplot') + 
  coord_cartesian(ylim = c(0,200))

by(pf$friendships_initiated, pf$gender, summary)
```
Response:

***

### Getting Logical
Notes:

```{r Getting Logical}
pf$mobile_check_in <- NA
pf$mobile_check_in <- ifelse(pf$mobile_likes > 0, 1, 0)
pf$mobile_check_in <- factor(pf$mobile_check_in)
summary(pf$mobile_check_in)

percent_mobile_check_in <- sum(pf$mobile_check_in==1)/length(pf$mobile_check_in)
percent_mobile_check_in
```

Response:

***

### Analyzing One Variable
Reflection on what I learned this lesson:
- Syntax for 'R'
- More about box plots
- Logarithmic scaling
- Manipulating graphs in R (scale_x_continuous, breaks, limits, facet_wraps)
***

Click **KnitHTML** to see all of your hard work and to have an html
page of this lesson, your answers, and your notes!