Lesson 4
========================================================

***

### Scatterplots and Perceived Audience Size
Notes:
```{r}
library(ggplot2)
pf <- read.csv("/Users/markmeleka/Documents/Comp Sci and Programming Learning/Udacity/DAND/Exploratory_Data_Analysis/eda-course-materials/lesson3/pseudo_facebook.tsv", sep = "\t")

setwd("/Users/markmeleka/Documents/Comp Sci and Programming Learning/Udacity/DAND/Exploratory_Data_Analysis/eda-course-materials/lesson4")
```

***

### Scatterplots
Notes:

```{r Scatterplots}
qplot(x=age, y= friend_count, data = pf) + 
  scale_x_continuous(breaks=seq(10,115,5))
```

***

#### What are some things that you notice right away?
Response:

***

### ggplot Syntax
Notes:

```{r ggplot Syntax}
ggplot(aes(x=age, y= friend_count), data = pf) + 
  geom_point() +
  xlim(13,90)
```

***

### Overplotting
Notes:

```{r Overplotting}
ggplot(aes(x=age, y= friend_count), data = pf) + 
  geom_jitter(alpha = 1/20) +
  xlim(13,90)
```

#### What do you notice in the plot?
Response:
Users under age 25 have the most instances of higher friend counts. There's still a spike around age.
***

### Coord_trans()
Notes:

```{r Coord_trans()}

```

#### Look up the documentation for coord_trans() and add a layer to the plot that transforms friend_count using the square root function. Create your plot!

```{r}
ggplot(aes(x=age, y= friend_count), data = pf) + 
  #must constrain 'jitter' in the y direction (using position=position_jitter(h=0) which tells 'geom_jitter()' not to 'jitter' in the height direction
  geom_jitter(alpha = 1/20, position=position_jitter(h=0)) +
  xlim(13,90) + 
  coord_trans(y= "sqrt")
```

#### What do you notice?

***

### Alpha and Jitter
Notes:

```{r Alpha and Jitter}
# Examine the relationship between
# friendships_initiated (y) and age (x)
# using the ggplot syntax.

ggplot(aes(x=age, y=friendships_initiated), data = pf) + 
  geom_jitter(alpha = 1/20, position=position_jitter(h=0)) + 
  coord_trans(y="sqrt")
```

***

### Overplotting and Domain Knowledge
Notes:

***

### Conditional Means
Notes:

```{r Conditional Means}
install.packages('dplyr', repos='https://cran.rstudio.com')
#had to add this because original glue link produced a 404.
#BUT, glue is not available for R v3.4.2
install.packages('glue', repos='https://cran.rstudio.com/bin/macosx/el-capitan/contrib/3.4/glue_1.2.0.tgz')

#let's try installing dplyr from a local path
#this didn't work
install.packages('dplyr', repos="/Users/markmeleka/Documents/Comp\ Sci\ and\ Programming\ Learning/Udacity/DAND/Exploratory_Data_Analysis/eda-course-materials/lesson4/lesson4_student.rmd")

#https://stackoverflow.com/questions/25721884/how-should-i-deal-with-package-xxx-is-not-available-for-r-version-x-y-z-wa
##6: downloaded through 'devtools' library + install 'glue' through github -> install 'dplyr' through github
install.packages("devtools")
library(devtools)
install_github("cran/glue")
install_github("cran/dplyr")


library(dplyr)

#commonly used functions:
filter()
group_by()
mutate()
arrange()

#an example
age_groups <- group_by(pf,age)
pf.fc_by_age <- summarise(age_groups,
                          friend_count_mean = mean(friend_count),
                          friend_count_median = median(friend_count),
                          n = n())
pf.fc_by_age <- arrange(pf.fc_by_age, age)

head(age_groups)
head(pf.fc_by_age)
```

```{R}
#same as the above example
pf.fc_by_age <- pf %>%
  group_by(age) %>%
  summarise(friend_count_mean = mean(friend_count),
            friend_count_median = median(friend_count),
            n = n()) %>%
  arrange(age)

head(pf.fc_by_age)
```

Create your plot!

```{r Conditional Means Plot}
ggplot(aes(x=age, y=friend_count_mean), data=pf.fc_by_age) + 
  geom_line() + 
  geom_point()
```

***

### Overlaying Summaries with Raw Data
Notes:

```{r Overlaying Summaries with Raw Data}
ggplot(aes(x=age, y=friend_count), data=pf) +
  geom_point(alpha=0.05,
             position=position_jitter(h=0),
             color="orange") + 
  coord_cartesian(xlim=c(13,70), ylim=c(0,1000)) +
  geom_line(stat = 'summary', fun.y =mean) + 
  geom_line(stat = 'summary', fun.y = quantile, fun.args= list(probs= 0.1),
            linetype=2, color='blue') +
  geom_line(stat = 'summary', fun.y = quantile, fun.args= list(probs= 0.9),
            linetype=2, color='blue') +
  geom_line(stat = 'summary', fun.y = quantile, fun.args= list(probs= 0.5),
            color='blue')
```

#### What are some of your observations of the plot?
Response:

***

### Moira: Histogram Summary and Scatterplot
See the Instructor Notes of this video to download Moira's paper on perceived audience size and to see the final plot.

Notes:

***

### Correlation
Notes:

```{r Correlation}
cor.test(pf$age, pf$friend_count)

#OR using with()
with(pf, cor.test(age, friend_count))
```

Look up the documentation for the cor.test function.

What's the correlation between age and friend count? Round to three decimal places.
Response:

***

### Correlation on Subsets
Notes:

```{r Correlation on Subsets}
with(subset(pf, age <= 70), cor.test(age, friend_count))
```

***

### Correlation Methods
Notes:

***

## Create Scatterplots
Notes:
# Create a scatterplot of likes_received (y)
# vs. www_likes_received (x).
```{r}
www_upper_quartile = quantile(pf$www_likes_received, probs=0.75)
likes_upper_quartile = quantile(pf$likes_received, probs=0.75)

ggplot(aes(x=www_likes_received, y=likes_received), data=pf) +
  geom_point() + 
  coord_cartesian(xlim=c(0, quantile(pf$www_likes_received, probs=0.95)), 
                  ylim=c(0,quantile(pf$likes_received, probs=0.95))) +
  geom_smooth(method='lm', color = 'red')

#I wanted a dynamically calculated var. for the x/y lim. The below didn't work though.
#www_upper_quartile+1.5*(IQR(pf$www_likes_recieved))

```

***

### Strong Correlations
Notes:

```{r Strong Correlations}
cor.test(pf$www_likes_received, pf$likes_received)
```

What's the correlation betwen the two variables? Include the top 5% of values for the variable in the calculation and round to 3 decimal places.

```{r Correlation Calcuation}

```

Response:

***

### Moira on Correlation
Notes:

***

### More Caution with Correlation
Notes:

```{r More Caution With Correlation}
install.packages('alr3')
library(alr3)

head(Mitchell)
```

Create your plot!

```{r Temp vs Month}
ggplot(aes(x=Month, y=Temp), data=Mitchell) +
  geom_point() + 
  geom_smooth()
```

***

### Noisy Scatterplots
a. Take a guess for the correlation coefficient for the scatterplot.

b. What is the actual correlation of the two variables?
(Round to the thousandths place)

```{r Noisy Scatterplots}
cor.test(Mitchell$Month, Mitchell$Temp)
```

***

### Making Sense of Data
Notes:

```{r Making Sense of Data}
ggplot(aes(x=Month, y=Temp), data=Mitchell) +
  geom_point() + 
  geom_smooth() + 
  scale_x_continuous(breaks=seq(0,203,12))
```

***

### A New Perspective

```{R}
ggplot(aes(x=(Month%%12),y=Temp), data=Mitchell)+
  geom_point() +
  geom_smooth()
```

```{R}
#using 'energy' package to for dcor.test: 
# Distance correlation t-test of multivariate independence.

install.packages('energy')
library(energy)
dcor.ttest(Mitchell$Month, Mitchell$Temp)

#testing example
x <- seq(0, 4*pi, pi/20)
y <- cos(x)
qplot(x = x, y = y)
dcor.ttest(x, y)
```

What do you notice?
Response:

Watch the solution video and check out the Instructor Notes!
Notes:

***

### Understanding Noise: Age to Age Months
Notes:

```{r Understanding Noise: Age to Age Months}
ggplot(aes(x=age, y=friend_count_mean), data= pf.fc_by_age) +
  geom_line() + 
  geom_smooth()

# Assume the reference date for calculating age is December 31, 2013.
pf$age_with_months <- (pf$age)+((12-pf$dob_month)/12)
```

***

### Age with Months Means

```{r Age with Months Means}

```

Programming Assignment
```{r Programming Assignment}
# Create a new data frame called
# pf.fc_by_age_months that contains
# the mean friend count, the median friend
# count, and the number of users in each
# group of age_with_months. The rows of the
# data frame should be arranged in increasing
# order by the age_with_months variable.

# For example, the first two rows of the resulting
# data frame would look something like...

# age_with_months  friend_count_mean	friend_count_median	n
#              13            275.0000                   275 2
#        13.25000            133.2000                   101 11

pf.fc_by_age_months <- pf %>%
  group_by(age_with_months) %>%
  summarise(friend_count_mean = mean(friend_count),
            friend_count_median = median(friend_count),
            n = n()) %>%
  arrange(age_with_months)
  
```

***

### Noise in Conditional Means

```{r Noise in Conditional Means}
ggplot(aes(x=age_with_months, y=friend_count_mean), data= subset(pf.fc_by_age_months, age_with_months < 71)) +
  geom_line() + 
  geom_smooth()
```

***

### Smoothing Conditional Means
Notes:

```{r Smoothing Conditional Means}

```

***

### Which Plot to Choose?
Notes:

***

### Analyzing Two Variables
Reflection:

***

Click **KnitHTML** to see all of your hard work and to have an html
page of this lesson, your answers, and your notes!

