---
title: "Exploring Prosper Loan Data"
author: Mark Meleka
date: February 10th, 2018
output: 
  html_document:
    toc: true
    toc_depth: 3
    toc_float: true
---

```{r echo=FALSE, message=FALSE, warning=FALSE, initial_packages}
library(ggplot2)
library(RColorBrewer)
```

```{r echo=FALSE, message=FALSE, warning=FALSE, Load_the_Data}
# Load the Data
loanData <- read.csv('prosperLoanData.csv')
# data source:
# https://s3.amazonaws.com/udacity-hosted-downloads/ud651/prosperLoanData.csv&sa=D&ust=1518297616552000&usg=AFQjCNHryWNnloE0Nd9Au1hj2OJqOrKolg

# variable definitions: 
# https://docs.google.com/spreadsheets/d/1gDyi_L4UvIrLTEC6Wri5nbaMmkGmLQBk-Yx3z0XDEtI/edit?usp=sharing

# creating effective plots:
# https://docs.google.com/document/d/1-f3wM3mJSkoWxDmPjsyRnWvNgM57YUPloucOIl07l4c/pub
```

#Intro 
This data set from Prosper, a peer-to-peer lending company, contains 113,937 
loans with 81 variables on each loan, including loan amount, borrower rate (or 
interest rate), current loan status, borrower income, borrower employment 
status, borrower credit history, and the latest payment information.

I chose this data set because I'm personally in the middle of researching 
mortgages and their associated rates. Mortgages and loans have a lot in common 
so I expect to learn a lot, even though the data set is from the United States 
and I live in Canada. Hopefully at the end of this exploration both myself and 
the reader will have a better understanding of what factors make one more likely 
to secure a larger loan at a low interest rate.

We have 113,937 observations of 81 variables (see Appendix 1). The goal of this 
report is to explore 10-15 variables. I'll start by getting a general intuition 
of the data. I noticed there isn't a variable for "Loan Amount Approved", but 
just a variable for the amount of the loan requested ("LoanOriginalAmount"). For 
this reason, I'll mainly be focusing on understanding the effect of different 
variables on borrower interest rates.

Variable definitions can be found here: 
https://docs.google.com/spreadsheets/d/1gDyi_L4UvIrLTEC6Wri5nbaMmkGmLQBk-Yx3z0XDEtI/edit?usp=sharing.

**Note: if you're short on time, I suggest skipping to the "Final Plots and 
Summary" section. If you have time after that, work your 
way backwards through the Multivariate exploration, and then to the Bivariate 
exploration.**

### Getting Basic Intuition About the Data

I looked at a summary of the data (see Appendix 2) and got a quick overview of 
each of the variables. These already give me some ideas of what I'm interested 
in exploring:  

* What's the relationship between Occupation and StatedMonthlyIncome?  
* StatedMonthlyIncome and LoanOriginalAmount?  
* StatedMonthlyIncome and CreditGrade?  
* StatedMonthlyIncome and CreditScore? ProsperScore?  
* What effect does being a home owner have? (IsBorrowerHomeowner)  
* BorrowerRate and all of the above?  

I wonder if we can make enough inferences to build a decent model for 
BorrowerRate later?

# Univariate Plots

```{r echo=FALSE, message=FALSE, warning=FALSE, Univariate_StatedMonthlyIncome}
qplot(loanData$StatedMonthlyIncome, xlim=c(0,30000))
summary(loanData$StatedMonthlyIncome)
loanData <- subset(loanData, StatedMonthlyIncome <= 100000)
```

Most people in our dataset make under $10,000 per month. 

Note that there are some StatedMonthlyIncome outliers that force us to limit the 
x-axis scale. I also removed data points where StatedMonthlyIncome was greater than 100000. 
This was because 99% of the data was lower than this value, and these outliers 
were making it difficult to garner useful insights from the data.

```{r echo=FALSE, message=FALSE, warning=FALSE, Univariate_EmploymentStatus}
# Re-ordering EmploymentStatus based on my gut feeling of which would qualify for a bigger loan.
loanData$EmploymentStatus <- ordered(loanData$EmploymentStatus, 
                                     levels=c("Employed", 
                                              "Full-time", 
                                              "Part-time", 
                                              "Self-employed", 
                                              "Retired", 
                                              "Not employed", 
                                              "Not available", 
                                              "Other", 
                                              ""))

ggplot(loanData,
       aes(x=EmploymentStatus)) + 
  geom_bar() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))

summary(loanData$EmploymentStatus)
```

Our dataset contains many people who are employed or working full-time. I'm not 
sure the difference between 'Employed', 'Full-time' and 'Part-time'. It seems 
like the options aren't mutually exclusive.

```{r echo=FALSE, message=FALSE, warning=FALSE, Univariate_LoanOriginalAmount}
qplot(loanData$LoanOriginalAmount)
summary(loanData$LoanOriginalAmount)
```

Most loans are under \$20,000. It looks like there are several spikes in $5000 
increments. This might be simply because they're simple round numbers.

```{r echo=FALSE, message=FALSE, warning=FALSE, Univariate_CreditGrade}
# Ordering CreditGrade based on the actual heirarchy of credit grades.
loanData$CreditGrade <- ordered(loanData$CreditGrade, 
                                levels=c('AA', 
                                         'A', 
                                         'B', 
                                         'C', 
                                         'D', 
                                         'E', 
                                         'HR', 
                                         'NC', 
                                         ''))

qplot(loanData$CreditGrade)
summary(loanData$CreditGrade)

```

It seems that most of our loans don't have a credit grade assigned.

```{r echo=FALSE, message=FALSE, warning=FALSE, Univariate_LoanStatus}
ggplot(loanData,
       aes(x=LoanStatus)) + 
  geom_bar() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))

summary(loanData$LoanStatus)
```

It's hard to compare Past Due loans because they're distributed by number of 
months past due. Let's group them all together for easy trending.

```{r echo=FALSE, message=FALSE, warning=FALSE, Univariate_LoanStatusGeneral}
loanData$LoanStatusGeneral <- loanData$LoanStatus

levels(loanData$LoanStatusGeneral) <- c(levels(loanData$LoanStatusGeneral), 
                                        'PastDue')

past_due_rows <- grep('^Past', loanData$LoanStatusGeneral)

loanData$LoanStatusGeneral[past_due_rows] <- 'PastDue'

loanData$LoanStatusGeneral <- ordered(loanData$LoanStatusGeneral, 
                                      levels=c("Completed", 
                                               "FinalPaymentInProgress", 
                                               "Current", 
                                               "PastDue", 
                                               "Defaulted", 
                                               "Chargedoff", 
                                               "Cancelled"))

ggplot(loanData,
       aes(x=LoanStatusGeneral)) + 
  geom_bar() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))

summary(loanData$LoanStatusGeneral)
```

Most loans are either Current, Completed, or ChargedOff.

Note: Here's are definitions for the various loan statuses:
http://help.bitbond.com/article/20-the-10-loan-status-variants-explained

```{r echo=FALSE, message=FALSE, warning=FALSE, Univariate_CreditScoreAverage}
loanData$CreditScoreAverage <- (loanData$CreditScoreRangeLower +
                                  loanData$CreditScoreRangeUpper) /2

qplot(loanData$CreditScoreAverage, 
      xlim=c(400,900), 
      bins=25)

summary(loanData$CreditScoreAverage)
```

I made a new variable called 'CreditScoreAverage' which is the average of 
CreditScoreRangeLower and CreditScoreRangeUpper Most average credit scores are 
between 600 and 800.

```{r echo=FALSE, message=FALSE, warning=FALSE, Univariate_BorrowerRate}
qplot(loanData$BorrowerRate)
summary(loanData$BorrowerRate)

```

Most of the borrower interest rates are between 0.05 and 0.35. Unfortunately, 
the interest rate units are missing. It seems unlikely that interest rates are 
between 5% to 35% or 0.05% to 0.35%. It seems far more likely that they're 
between 0.5% to 3.5%.

```{r echo=FALSE, message=FALSE, warning=FALSE, Univariate_ListingYear}
loanData$ListingYear <- substring(loanData$ListingCreationDate,1,4)

qplot(loanData$ListingYear)
```

Most of the data is between 2006 and 2014. We see a sharp drop of data in 2009. 
This is likely due to the 2008 financial crisis and ensuing recession. After 
this, we see year-over-year growth in the number of Prosper loans, until 2014. 
This is likely because the dataset was distributed in 2014 and this year might 
include incomplete data.

```{r echo=FALSE, message=FALSE, warning=FALSE, Univariate_2014Loans}
ggplot(subset(loanData, ListingYear == 2014),
       aes(x= LoanOriginationQuarter)) + 
  geom_bar()
```

Here we confirm we do have incomplete data for 2014. We only have data from Q1.

# Univariate Analysis
### Dataset Structure
The original dataset contained 113,937 observations of 81 variables. Some of these 
variables include:

* loan amount (mostly under $20k),
* credit grade of the loan (mostly missing),
* loan status (mainly completed, current, or defaulted),
* borrower credit score (mainly between 600 and 800), 
* borrower interest rate (mainly between 0.05 and 0.35), and
* listing year (mainly 2006-2013).

### Main Features of Interest
I'm curious how credit score impacts loan amount and borrower interest rate. 
I'll likely want to get into the details of the data so I'll facet by year.

### Other Potentially Interesting Features
I wonder if having one's friends invest in you increases the likelihood of a 
successful loan (ie. does InvestmentFromFriendsCount affect LoanStatus)? I also 
wonder how much money Prosper made from 2009 to 2013, since the number of loans 
those years kept increasing year-over-year.

### New Variables Created
I created two new variables: 

* ListingYear (a function of ListingCreationDate), and
* CreditScoreAverage (a function of CreditScoreRangeLower and 
 CreditScoreRangeUpper).

### Interesting Distributions
The StatedMonthlyIncome distribution was right-skewed with a mean of \$5608. 
There seemed to be an outlier(s) since the max monthly income was $1,750,003. 
CreditScoreAverage and BorrowerRate weer noramlly distributed with a mean of 695 
and 0.19 respectively.

I chose to re-order EmploymentStatus, CreditGrade, and LoanStatus so that the 
plots would have an apparent natural order. 

# Bivariate Plots
The first thing I'll do is get a quick high-level overview of which interesting 
numeric values are correlated.

```{r echo=FALSE, message=FALSE, warning=FALSE, Bivariate_Pairs_plot}
library(magrittr)
#install.packages('ggcorrplot')
library(ggcorrplot)

cor_matric <- cor(loanData[ , c('StatedMonthlyIncome', 
                  'CreditScoreAverage', 
                  'ProsperScore',
                  'BorrowerRate',
                  'EstimatedReturn',
                  'OpenCreditLines',
                  'TotalCreditLinespast7years',
                  'InquiriesLast6Months',
                  'CurrentDelinquencies',
                  'LoanOriginalAmount',
                  'DebtToIncomeRatio')], 
    use="complete.obs")

ggcorrplot(cor_matric, 
           hc.order = TRUE, 
           type = "lower", 
           lab = TRUE, 
           lab_size = 3, 
           method="circle", 
           title="Correlogram of loanData quant variables", 
           ggtheme=theme_bw)
```

Here we see a quick overview of how the chosen quantitative variables relate to 
each other. We can see that the variable pairs with a strong correlation are:
- BorrowerRate and EstimatedReturn (0.83)
- ProsperScore and BorrowerRate (-0.66)
- OpenCreditLines and TotalCreditLinespast7years (0.57) (although this seems 
 obvious)
- CreditScoreAverage and BorrowerRate (-0.53)

But first, I'm curious which jobs make the most money! 

```{R echo=FALSE, message=FALSE, warning=FALSE, Bivariate_StatedMonthlyIncome_Occupation}
#Focusing only on verifiable incomes, which represent more than 90% of the data.
ggplot(subset(loanData, IncomeVerifiable = 'True'), 
       aes(x= reorder(Occupation, 
                      StatedMonthlyIncome, 
                      FUN=median), 
           y=StatedMonthlyIncome)) +
  geom_boxplot() + 
  coord_flip(ylim= c(0, 30000)) + 
  theme(axis.text.y=element_text(size=5))

#median(loanData$StatedMonthlyIncome[loanData$Occupation == 'Analyst'])
```

Doctors, judges, and pharmacists make the most money, with median incomes 
greater than \$10,000 per month. The median salary for an analyst is only 
\$5800 per month. Is it too late to switch careers? 

I'm suprised that in our quick correlation analysis above, LoanOriginalAmount 
and StatedMonthlyIncome didn't have a strong positive relation. I want to see 
this visually.

```{r echo=FALSE, message=FALSE, warning=FALSE, Bivariate_StatedMonthlyIncome_LoanOriginalAmount}
ggplot(loanData, aes(x=StatedMonthlyIncome, y=LoanOriginalAmount)) + 
  geom_jitter(alpha = 0.1) + 
  geom_smooth() +
  coord_cartesian(xlim = c(0, 30000), 
                  ylim = c(0, max(loanData$LoanOriginalAmount)))
```

So we do see a moderate positive relation, at least when StatedMonthlyIncome is 
between \$2000 and \$10,000. After that, increasing income doesn't seem to have 
a strong impact on the amount of the loans asked for. Note that this isn't the 
amount of loan the borrower is approved for, it's just how much they asked for.

Let's look at StatedMonthlyIncome and BorrowerRate.

```{R echo=FALSE, message=FALSE, warning=FALSE, Bivariate_StatedMonthlyIncome_BorrowerRate}
ggplot(subset(loanData, StatedMonthlyIncome < 25000),
       aes(x= StatedMonthlyIncome,
           y=BorrowerRate)) + 
  geom_jitter(alpha=0.1) + 
  geom_smooth()
```

We don't see a visual correlation above.

```{R echo=FALSE, message=FALSE, warning=FALSE, Bivariate_CreditScoreAverage_BorrowerRate}
ggplot(loanData,
       aes(x= CreditScoreAverage,
           y=BorrowerRate)) + 
  geom_jitter(alpha=0.1) + 
  geom_smooth() + 
  coord_cartesian(xlim = c(450, 900))
```

It's clear that as credit score increases, BorrowerRate decreases.

What about ProsperScore? Is this basically a Prosper version of CreditScore?

```{R echo=FALSE, message=FALSE, warning=FALSE, Bivariate_ProsperScore_BorrowerRate}
ggplot(loanData,
       aes(x= ProsperScore,
           y=BorrowerRate)) + 
  geom_jitter(alpha=0.1) + 
  geom_smooth()
```

There's a similar trend line as the CreditScore graph: as ProsperScore increases 
BorrowerRate decreases.

```{R echo=FALSE, message=FALSE, warning=FALSE, Bivariate_LoanStatusGeneral_CreditScoreAverage}
ggplot(loanData,
       aes(x= LoanStatusGeneral,
           y=CreditScoreAverage)) + 
  geom_boxplot() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))

```

Looking at median CreditScores, we do see that lower credit scores have a higher 
chance of defaulting or charging off.

How much does BorrowerRate affect the estimated lender return.

```{R echo=FALSE, message=FALSE, warning=FALSE, Bivariate_BorrowerRate_EstimatedReturn}
ggplot(loanData,
       aes(x= BorrowerRate,
           y=EstimatedReturn)) + 
  geom_jitter(alpha=0.1) + 
  geom_smooth()
```

This makes sense. It's actually likely that EstimatedReturn is a function of 
BorrowerRate, although this isn't confirmed.

```{R echo=FALSE, message=FALSE, warning=FALSE, Bivariate_ListingYear_BorrowerRate}
ggplot(subset(loanData, ListingYear > 2005),
       aes(x= ListingYear,
           y=BorrowerRate)) + 
  geom_boxplot() +
  stat_summary(fun.y = "median", 
               geom = 'line', 
               group = 1, 
               size = 1, 
               color = 'red')
```

We can see Borrower Rates over time here. Note a dip in interest rates around 
the 2008 financial crisis.

```{R echo=FALSE, message=FALSE, warning=FALSE, Bivariate_ListingYear_LoanOriginalAmount}
ggplot(subset(loanData, ListingYear > 2005),
       aes(x= ListingYear,
           y=LoanOriginalAmount)) + 
  geom_boxplot() +
  stat_summary(fun.y = "median", 
               geom = 'line', 
               group = 1, 
               size = 1, 
               color = 'green')
```

Median loan amounts were between \$3000 and \$6000 until 2013 when they doubled 
to about $10,000.

```{R echo=FALSE, message=FALSE, warning=FALSE, Bivariate_ListingYear_LP_InterestandFees}
ggplot(loanData,
       aes(x=ListingYear,
           y=LP_InterestandFees)) + 
  stat_summary(fun.y = sum,
               geom = "bar")
```

Here we see the interest and fees Prosper collected by loan. We see that this 
number dropped off significantly in 2009 after the 2008 financial crisis. Then 
it continued to rise year-over-year until 2013. We don't have complete data for 
2014, so we can't make any inferenced about it.

# Bivariate Analysis
### Relationships Found With Main Features of Interest
There's a moderate correlation between LoanOriginalAmount and 
StatedMonthlyIncome when StatedMonthlyIncome is between \$2000 and $10,000. 
After that, increasing income doesn't seem to have a strong impact on the amount 
of the loans asked for. Note that this isn't the amount of loan the borrower is 
approved for, it's just how much they asked for.

I didn't see a visual correlation between StatedMonthlyIncome and BorrowerRate.

We saw a correlation between CreditScore and ProsperScore. I plotted them both 
against BorrowerRate and saw a similar trend line (as CreditScore or 
ProsperScore increased, BorrowerRate decreased).

Additionally, we saw a strong negative correlation between CreditScore and 
BorrowerRate. 

### Other Interesting Relationships
Looking at StatedMonthlyIncome and Occupation we see that doctors, judges, and 
pharmacists make the most money, with median incomes greater than \$10,000 per 
month. 

Lower CreditScores had a higher chance of defaulting or charging off compared 
to higher ones.

BorrowerRate is highly correlated to estimated lender return. One is likely a 
function of the other.

When looking at BorrowerRates over time, we see an expected dip in interest 
rates around the 2008 financial crisis.

Looking at loan amounts, we see that they were between \$3000 and \$6000 until 
2013 and then doubled to about $10,000.

### Strongest Relationships Found
The strongest relationship I found was between BorrowerRate and EstimatedReturn. 
Although one is likely a function of the other. The second highest was between 
ProsperScore and BorrowerRate.

# Multivariate Plots
Let's start by looking at MonthlyIncome vs. LoanAmount vs. LoanStatus to get an 
overview.

```{r echo=FALSE, message=FALSE, warning=FALSE, Multivariate_StatedMonthlyIncome_LoanOriginalAmount_LoanStatusGeneral}
#install.packages('ggalt')
library(ggalt)

ggplot(loanData, aes(x=StatedMonthlyIncome, 
                     y=LoanOriginalAmount, 
                     color=LoanStatusGeneral)) + 
  geom_jitter(alpha = 0.05) + 
  scale_color_brewer(type = 'div',
                     guide = guide_legend(title = 'Loan Status',
                                          override.aes = list(alpha = 1, 
                                                              size = 2)),
                     direction = -1) +
  coord_cartesian(xlim = c(0, 30000), 
                  ylim = c(0, max(loanData$LoanOriginalAmount))) + 
  geom_encircle() + 
  theme_dark()
```

Here we see many 'completed' loans where the LoanOriginalAmount is low. 
Defaulted and ChargedOff loans seem to occur when the StatedMonthlyIncome is 
low and the LoanOriginalAmount is high. It's also clear that loans only get 
cancelled when monthly income and loan amount is lower than $5000.

Let's look at a similar graph, but colour the data by EmploymentStatus.

```{r echo=FALSE, message=FALSE, warning=FALSE, Multivariate_StatedMonthlyIncome_LoanOriginalAmount_EmploymentStatus}
ggplot(subset(loanData, EmploymentStatus != ''), 
       aes(x=StatedMonthlyIncome,
           y=LoanOriginalAmount,
           color=EmploymentStatus)) + 
  geom_jitter(alpha = 0.05) + 
  scale_color_brewer(type = 'div',
                     palette = 4,
                     guide = guide_legend(title = 'Employment Status',
                                          override.aes = list(alpha = 1, 
                                                              size = 2)),
                     direction = -1) +
  #geom_smooth() +
  coord_cartesian(xlim = c(0,30000), 
                  ylim = c(0,max(loanData$LoanOriginalAmount)))
```

Hard to see more than the obvious: retired/not employed/not available workers 
have about 0 income.

I'm curious what kind of loans the employed group requested.

```{r echo=FALSE, message=FALSE, warning=FALSE, Multivariate_StatedMonthlyIncome_LoanOriginalAmount_EmploymentStatus2}
ggplot(subset(loanData, EmploymentStatus %in% c('Full-time', 
                                                'Part-time', 
                                                'Self-employed')), 
       aes(x=StatedMonthlyIncome,
           y=LoanOriginalAmount,
           color=EmploymentStatus)) + 
  geom_jitter(alpha = 0.1) + 
  scale_color_brewer(type = 'qual',
                     #palette = 1,
                     guide = guide_legend(title = 'Employment Status',
                                          override.aes = list(alpha = 1, 
                                                              size = 2)),
                     direction = -1) +
  #geom_smooth() +
  coord_cartesian(xlim = c(0,30000), 
                  ylim = c(0,max(loanData$LoanOriginalAmount))) + 
  theme_dark()
```

Part-time workers mostly made under \$2000/month, and didn't seem to ask for 
more than \$10000. Self-employed people seemed to mainly ask for \$4000, 
\$10,000 or \$15,000.

```{r echo=FALSE, message=FALSE, warning=FALSE, Multivariate_EmploymentStatus_LoanStatusGeneral_ListingYear}
ggplot(subset(loanData, EmploymentStatus != ''), 
       aes(x=EmploymentStatus,
           y=LoanStatusGeneral)) + 
  geom_jitter(alpha = 0.1) + 
  facet_wrap(~ListingYear) + 
  theme(axis.text.x = element_text(angle = 90, hjust = 1))
```

Using this graph we can immediately see the changes over time to how Prosper 
collected data. For example, we can see that they didn't keep EmploymentStatus 
data from 2006. From 2007 to 2009 they didn't have a 'Past Due', 'Current', or 
'FinalPaymentInProgress' loan status. 

We see very few loans in 2009, after the 2008 financial crisis. 2010 picks back 
up at a bit. In 2011 the graph is dense again, with many self-employed loans.

```{r echo=FALSE, message=FALSE, warning=FALSE, Multivariate_CreditScoreAverage_BorrowerRate_LoanOriginalAmount_ListingYear}
ggplot(loanData,
       aes(x= CreditScoreAverage,
           y=BorrowerRate)) + 
  geom_jitter(alpha=0.1, aes(color = LoanOriginalAmount)) + 
  scale_colour_gradientn(colours=rainbow(4)) +
  geom_smooth() + 
  coord_cartesian(xlim=c(500,850)) + 
  facet_wrap(~ListingYear)
```

Here we see the progression of Prosper's users over time. It starts with users 
with a wide range of credit scores. Over the years, the minimum credit score 
increases. Loan amounts in 2006 were mainly under \$10,000. We see these rise 
over time, until they're mainly \$15,000+ loans in 2013.

```{R echo=FALSE, message=FALSE, warning=FALSE, Multivariate_ListingYear_LP_InterestandFees}
loanData$IncomeRange <- ordered(loanData$IncomeRange, 
                                levels=c("Not displayed", 
                                         "Not employed", 
                                         "$0", "$1-24,999",  
                                         "$25,000-49,999", 
                                         "$50,000-74,999", 
                                         "$75,000-99,999", 
                                         "$100,000+"))

ggplot(subset(loanData, IncomeRange != "Not displayed"),
       aes(x=ListingYear,
           y=LP_InterestandFees/1000000,
           fill = IncomeRange)) + 
  scale_fill_brewer('div',
                    palette = 2,
                    guide = guide_legend(title = 'Income Range')) +
  stat_summary(fun.y = sum,
               geom = "bar",
               position = 'stack')
```

Since the 2008 financial crisis, Prosper has made more and more of it's money 
from higher income individuals.

```{r echo=FALSE, message=FALSE, warning=FALSE, LinearModel_BorrowerRate}
library(memisc)

br_m1 <- lm(BorrowerRate ~ ProsperScore, data = loanData)
br_m2 <- update(br_m1, ~ . + CreditScoreAverage)
br_m3 <- update(br_m2, ~ . + ListingYear)

mtable(br_m1, br_m2, br_m3, sdigits = 3)

```

We've made a model here to try to explain the variance in borrower rate. It 
looks like it's mainly made up of ProsperScore/CreditScore and ListingYear. I 
tried adding a few other variables (StatedMonthlyIncome, Term, 
IsBorrowerHomeowner), but they either didn't have a significant effect on 
explaining the variance in BorrowerRate, or they had a negative effect.

# Multivariate Analysis
### New Discoveries in the Data
We saw a positive relationship between loan amount and year of the loan. Other 
than this, we didn't discover any new relationships, but did learn about the 
nuances of previously discovered relationships. For example: 

* Many 'completed' loans where the LoanOriginalAmount is low.  
* Defaulted and ChargedOff loans seem to occur when the StatedMonthlyIncome is 
 low and the LoanOriginalAmount is high.  
* Part-time workers mostly made under \$2000/month, and didn't seem to ask for 
 more than \$10000.  
* Self-employed people seemed to mainly ask for \$4000, \$10,000 or \$15,000.  
* Prosper didn't keep EmploymentStatus data from 2006. From 2007 to 2009 they 
 didn't have a 'Past Due', 'Current', or 'FinalPaymentInProgress' loan status.   
* We see very few loans in 2009, after the 2008 financial crisis. 2010 picks 
 back up at a bit. In 2011 the graph is dense again, with many self-employed 
 loans.  
* In earlier years Prosper users have a wide range of credit scores. Over the 
 years, the minimum credit score increases. Loan amounts in 2006 were mainly 
 under \$10,000. We see these rise over time, until they're mainly \$15,000+ 
 loans in 2013.  
* Since the 2008 financial crisis, Prosper has made more and more of it's money 
 from higher income individuals.  

### Linear Models
I made a model to try to explain the variance in borrower rate. It used 
ProsperScore/CreditScore and ListingYear and explained 68% of the variance in 
borrower rate. I tried adding a few other variables (StatedMonthlyIncome, Term, 
IsBorrowerHomeowner), but they either didn't have a significant effect on 
explaining the variance in BorrowerRate, or they had a negative effect. 

------

# Final Plots and Summary
### Plot One: Correlation Pairs
```{r echo=FALSE, message=FALSE, warning=FALSE, FinalPlot_Pairs_plot}
cor_matric <- cor(loanData[ , c('StatedMonthlyIncome', 
                  'CreditScoreAverage', 
                  'ProsperScore',
                  'BorrowerRate',
                  'EstimatedReturn',
                  'OpenCreditLines',
                  'TotalCreditLinespast7years',
                  'InquiriesLast6Months',
                  'CurrentDelinquencies',
                  'LoanOriginalAmount',
                  'DebtToIncomeRatio'
                  )], 
    use="complete.obs")

ggcorrplot(cor_matric, 
           hc.order = TRUE, 
           type = "lower", 
           lab = TRUE, 
           lab_size = 3, 
           method="circle", 
           title="Correlogram of quantitative loanData variables", 
           legend.title = "Correlation",
           ggtheme=theme_bw)
```

This plot shows the reader a quick and easy overview of how the chosen 
quantitative variables relate to each other. Using it, we can see that the 
variable pairs with a strong correlation are:

* BorrowerRate and EstimatedReturn (0.83)  
* ProsperScore and BorrowerRate (-0.66)  
* OpenCreditLines and TotalCreditLinespast7years (0.57) (although this seems 
 obvious)  
* CreditScoreAverage and BorrowerRate (-0.53)  

I used the above to give me a quick idea of which variables to dig deeper into.

### Plot Two: Prosper User Evolution
```{r echo=FALSE, message=FALSE, warning=FALSE, FinalPlot_CreditScoreAverage_BorrowerRate_LoanOriginalAmount_ListingYear}
ggplot(subset(loanData, ListingYear > 2005),
       aes(x= CreditScoreAverage,
           y=BorrowerRate/10)) + 
  geom_jitter(alpha=0.1, aes(color = LoanOriginalAmount)) + 
  scale_colour_gradientn(colours=rainbow(4),
                         name = 'Loan Amount ($)') +
  geom_smooth() + 
  coord_cartesian(xlim=c(500,850)) + 
  facet_wrap(~ListingYear) + 
  ggtitle("Borrower Rate by Credit Score, 
          showing loan amount, faceted by year") + 
  labs(x="Average Credit Score", y= "Borrower Interest Rate") + 
  scale_y_continuous(labels = scales::percent) + 
  guides(fill=guide_legend(title="Loan Amount"))

```

This plot shows the progression of Prosper's users (or Prosper's focus) over 
time. We can see that Prosper in 2006 had users with a wide range of credit 
scores. Over the years, the minimum credit score increased. We also see that 
loan amounts in 2006 were mainly under \$10,000 and we see these rise over time, 
until the majority of loans in 2013 are above \$15,000.


### Plot Three: Interest and Fees Collected
```{R echo=FALSE, message=FALSE, warning=FALSE, FinalPlot_ListingYear_LP_InterestandFees_IncomeRange}
library(scales)

ggplot(subset(loanData, IncomeRange != "Not displayed"),
       aes(x=ListingYear,
           y=LP_InterestandFees/1000000,
           fill = IncomeRange)) + 
  scale_fill_brewer('div',
                    palette = 2,
                    guide = guide_legend(title = "Borrower Income Range")) +
  stat_summary(fun.y = sum,
               geom = "bar",
               position = 'stack') + 
  ggtitle("Interest and Fees income by Year, 
          split by borrower's income range") + 
  labs(x="Listing Year", y= "Interest and Fees (millions)") + 
  scale_y_continuous(label= dollar_format())
```

It was important to me to also understand how Prosper performed financially as 
a company. If they did well, then I could expect other lenders to have similar 
criteria for choosing borrower interest rates. With this plot, we see that since 
the 2008 financial crisis, Prosper made more and more money from interest rates 
and fees each year. In particular, we see that more and more of this income came 
from higher income individuals.

------

# Final Reflection
I started this exploration so I could have a better understanding of what 
variables help one borrow more money with a low interest rate. I wasn't able to 
see which factors lead to being able to borrow more money, because there wasn't 
a "maximum loan amount approved for" variable. Instead, just a loan amount 
requested variable. This is fine though, I focused more of my efforts on 
borrower rate instead. 

I learned that the biggest factor to getting a lower interest rate is one's 
credit score, or the Prosper equivalent. I also learned that depending on the 
year, interest rates rise and fall. This is probably not something one can 
predict in advance, just like one can't predict a recession in advance. 

I also learned that with a data set with so many variables like this, it's easy 
to find oneself going down a rabbit hole looking at more and more relationships 
between variables. Several times I had to stop myself for the sake of time and 
not overloading the reader.

When I was stuck trying to figure out how to do something particular, I found 
it very helpful to look through the documentation or to do lots of Googling. It 
seemed like all the problems I went through, someone else online also went 
through before.

If I was to spend more time exploring this data set, I'd look for more 
relationships between other variables. In particular, I'd look to see what 
variables result in a higher credit score. I didn't look to deeply at credit 
lines, revolving accounts, inquiries, delinquincies, or debt-to-income ratios. 
I'd be interested in creating a model for credit score or Prosper score, but 
alass, I'll have to leave this for the next time. 

# Appendices
### Appendix 1: Data Structure
```{r echo=FALSE, message=FALSE, warning=FALSE, Data_structure}
str(loanData)
```

### Appendix 2: Data Summary
```{r echo=FALSE, message=FALSE, warning=FALSE, Data_summary}
summary(loanData)
```