Lesson 5
========================================================

### Multivariate Data
Notes:
```{R}
pf <- read.delim('/Users/markmeleka/Documents/Comp\ Sci\ and\ Programming\ Learning/Udacity/DAND/Exploratory_Data_Analysis/eda-course-materials/lesson3/pseudo_facebook.tsv')
library(dplyr)
```
***

### Moira Perceived Audience Size Colored by Age
Notes:

***

### Third Qualitative Variable
Notes:

```{r Third Qualitative Variable}
ggplot(aes(x = gender, y = age),
       data = subset(pf, !is.na(gender))) + geom_histogram()

# The data frame should contain the following variables:
#    mean_friend_count,
#    median_friend_count,
#    n (the number of users in each age and gender grouping)

pf.fc_by_age_gender <- pf %>%
  filter(!is.na(gender)) %>%
  group_by(age, gender) %>%
  summarise(mean_friend_count = mean(friend_count),
            median_friend_count = median(friend_count),
            n = n()) 

head(pf.fc_by_age_gender)

```

***

### Plotting Conditional Summaries
Notes:

```{r Plotting Conditional Summaries}
# Create a line graph showing the
# median friend count over the ages
# for each gender.

ggplot(aes(x=age, y=median_friend_count, color=gender), 
       data=pf.fc_by_age_gender) + 
  geom_line()

#OR

ggplot(aes(x=age, y=median_friend_count),
       data= pf.fc_by_age_gender) + 
  geom_line(aes(color=gender))

```

***

### Thinking in Ratios
Notes:

***

### Wide and Long Format
Notes:

***

### Reshaping Data
Notes:

```{r}
install.packages('reshape2')
library(reshape2)
#an intro to reshape2: http://seananderson.ca/2013/10/19/reshape.html

install.packages('tidyr')
library(tidyr)

#converting between wide and long format: http://www.cookbook-r.com/Manipulating_data/Converting_data_between_wide_and_long_format/

#melt data frames: http://www.r-bloggers.com/melt/

head(pf.fc_by_age_gender)
pf.fc_by_age_gender.wide <- dcast(pf.fc_by_age_gender,
                                  age ~ gender,
                                  value.var = 'mean_friend_count')
head(pf.fc_by_age_gender.wide)
```


***

### Ratio Plot
Notes:
# Plot the ratio of the female to male median
# friend counts using the data frame
# pf.fc_by_age_gender.wide.

# Think about what geom you should use.
# Add a horizontal line to the plot with
# a y intercept of 1, which will be the
# base line. Look up the documentation
# for geom_hline to do that. Use the parameter
# linetype in geom_hline to make the
# line dashed.

# The linetype parameter can take the values 0-6:
# 0 = blank, 1 = solid, 2 = dashed
# 3 = dotted, 4 = dotdash, 5 = longdash
# 6 = twodash

# This assignment is not graded and
# will be marked as correct when you submit.

# ENTER YOUR CODE BELOW THIS LINE
# =================================================
```{r Ratio Plot}
ggplot(aes(x= age, y = female/male), data=pf.fc_by_age_gender.wide) + 
  geom_line() + 
  geom_hline(yintercept=1, alpha = 0.3, linetype = 2)
```

***

### Third Quantitative Variable
Notes:

```{r Third Quantitative Variable}
pf$year_joined <- floor(2014 - pf$tenure/365)
```

***

### Cut a Variable
Notes:

```{r Cut a Variable}
summary(pf$year_joined)

ggplot(aes(x = year_joined), data= pf) +
  geom_bar()

# Create a new variable in the data frame
# called year_joined.bucket by using
# the cut function on the variable year_joined.

# You need to create the following buckets for the
# new variable, year_joined.bucket

#        (2004, 2009]
#        (2009, 2011]
#        (2011, 2012]
#        (2012, 2014]

# Note that a parenthesis means exclude the year and a
# bracket means include the year.

pf$year_joined.bucket <- cut(pf$year_joined, breaks = c(2004,2009,2011,2012,2014), right = TRUE)

table(pf$year_joined.bucket)
```

***

### Plotting it All Together
Notes:
# Create a line graph of friend_count vs. age
# so that each year_joined.bucket is a line
# tracking the median user friend_count across
# age. This means you should have four different
# lines on your plot.

# You should subset the data to exclude the users
# whose year_joined.bucket is NA.
```{r Plotting it All Together}
ggplot(aes(x= age, y= friend_count), 
       data= subset(pf, !is.na(year_joined.bucket))) +
  geom_line(aes(color = year_joined.bucket), 
            stat= 'summary', 
            fun.y = median)
```

***

### Plot the Grand Mean
Notes:
# Write code to do the following:

# (1) Add another geom_line to code below
# to plot the grand mean of the friend count vs age.

# (2) Exclude any users whose year_joined.bucket is NA.

# (3) Use a different line type for the grand mean.

# As a reminder, the parameter linetype can take the values 0-6:

# 0 = blank, 1 = solid, 2 = dashed
# 3 = dotted, 4 = dotdash, 5 = longdash
# 6 = twodash
```{r Plot the Grand Mean}
ggplot(aes(x= age, y= friend_count), 
       data= subset(pf, !is.na(year_joined.bucket))) +
  geom_line(aes(color = year_joined.bucket), 
            stat= 'summary', 
            fun.y = mean) + 
  geom_line(stat='summary', 
            fun.y = mean, 
            linetype = 2)
```

***

### Friending Rate
Notes:

```{r Friending Rate}
with(subset(pf, tenure>=1), summary(friend_count/tenure))
```

***

### Friendships Initiated
Notes:
# Create a line graph of mean of friendships_initiated per day (of tenure)
# vs. tenure colored by year_joined.bucket.

# You need to make use of the variables tenure,
# friendships_initiated, and year_joined.bucket.

# You also need to subset the data to only consider user with at least
# one day of tenure.
What is the median friend rate?
0.22
What is the maximum friend rate?
417
```{r Friendships Initiated}
ggplot(aes(x=7*round(tenure/7), y=friendships_initiated/tenure), data = subset(pf, tenure>=1)) + 
  geom_line(aes(color = year_joined.bucket), 
            stat='summary', 
            y.fun=mean)
```

***

### Bias-Variance Tradeoff Revisited
Notes:

```{r Bias-Variance Tradeoff Revisited}

ggplot(aes(x = tenure, y = friendships_initiated / tenure),
       data = subset(pf, tenure >= 1)) +
  geom_line(aes(color = year_joined.bucket),
            stat = 'summary',
            fun.y = mean)

ggplot(aes(x = 7 * round(tenure / 7), y = friendships_initiated / tenure),
       data = subset(pf, tenure > 0)) +
  geom_line(aes(color = year_joined.bucket),
            stat = "summary",
            fun.y = mean)

ggplot(aes(x = 30 * round(tenure / 30), y = friendships_initiated / tenure),
       data = subset(pf, tenure > 0)) +
  geom_line(aes(color = year_joined.bucket),
            stat = "summary",
            fun.y = mean)

ggplot(aes(x = 90 * round(tenure / 90), y = friendships_initiated / tenure),
       data = subset(pf, tenure > 0)) +
  geom_line(aes(color = year_joined.bucket),
            stat = "summary",
            fun.y = mean)

ggplot(aes(x = tenure, y = friendships_initiated / tenure),
       data = subset(pf, tenure >= 1)) +
  geom_smooth(aes(color = year_joined.bucket))

```

***

### Sean's NFL Fan Sentiment Study
Notes:

***

### Introducing the Yogurt Data Set
Notes:

***

### Histograms Revisited
Notes:

```{r Histograms Revisited}
yo <- read.csv('yogurt.csv')
str(yo)

yo$id <- factor(yo$id)
str(yo)

ggplot(aes(x = price), data=yo) + 
  geom_histogram(binwidth = 10)
```

***

### Number of Purchases
Notes:
# Create a new variable called all.purchases,
# which gives the total counts of yogurt for
# each observation or household.

# One way to do this is using the transform
# function. You can look up the function transform
# and run the examples of code at the bottom of the
# documentation to figure out what it does.

# The transform function produces a data frame
# so if you use it then save the result to 'yo'!
```{r Number of Purchases}
summary(yo$price)
length(unique(yo$price))
table(yo$price)

yo <- transform(yo, all.purchases= (strawberry + blueberry + pina.colada + plain + mixed.berry))
```

***

### Prices over Time
Notes:
# Create a scatterplot of price vs time.

# This will be an example of a time series plot.

# Resolve overplotting issues by using
# techniques you learned in Lesson 4.

# What are some things that you notice?
```{r Prices over Time}
ggplot(aes(x = time, y=price), data = yo) + 
  geom_jitter(alpha = 1/4 , shape = 21, fill = 'orange') + 
  geom_smooth()
```

***

### Sampling Observations
Notes:

***

### Looking at Samples of Households

```{r Looking at Sample of Households}
set.seed(4230)
sample.ids <- sample(levels(yo$id), 16)

ggplot(aes(x=time, y=price),
       data = subset(yo, id %in% sample.ids)) + 
  facet_wrap( ~ id) + 
  geom_line() + 
  geom_point(aes(size= all.purchases), pch = 1)
```

***

### The Limits of Cross Sectional Data
Notes:

***

### Many Variables
Notes:

***

### Scatterplot Matrix
Notes:
```{r Looking at Sample of Households}
install.packages('GGally')
library(GGally)
theme_set(theme_minimal(20))

set.seed(1836)
pf_subset <- pf[, c(2:15)]
#same as
#pf_subset <- subset(pf, select = -c(userid, year_joined, year_joined_bucket))

names(pf_subset)
ggpairs(pf_subset[sample.int(nrow(pf_subset), 1000),])
#Scatterplots are below the diagonal, and categorical variables, like gender, create faceted histograms.
```
***

### Even More Variables
Notes:

***

### Heat Maps
Notes:

```{r}
nci <- read.table("nci.tsv")
colnames(nci) <- c(1:64)
```

```{r}
nci.long.samp <- melt(as.matrix(nci[1:200,]))
names(nci.long.samp) <- c("gene", "case", "value")
head(nci.long.samp)

ggplot(aes(y = gene, x = case, fill = value),
  data = nci.long.samp) +
  geom_tile() +
  scale_fill_gradientn(colours = colorRampPalette(c("blue", "red"))(100))

#Melt data frames in R (http://www.r-bloggers.com/melt/)
#Micro-array/Gene Expression Data (https://s3.amazonaws.com/udacity-hosted-downloads/ud651/nci.tsv)

```


***

### Analyzing Three of More Variables
Reflection:

***

Click **KnitHTML** to see all of your hard work and to have an html
page of this lesson, your answers, and your notes!

